---
title: "Coursera - Machine Learning Project"
author: "Rob Reid"
date: "October 2014"
output: html_document
---

Assumes caret, randomforest and e1071 packages have been installed
Assumes data files have already been downloaded

### Load Data###
```{r}
pmlTrain <- read.csv('pml-training.csv')
pmlTest <- read.csv('pml-testing.csv')
```

### Partition data sets### 
Half of the data into each set
```{r}
library(caret)
dataSet <- createDataPartition(pmlTrain$classe, p=.5, list=FALSE)
dataTraining <- pmlTrain[dataSet,]
dataTesting <- pmlTrain[-dataSet,]
```

###  Histogram - Classe values### 
```{r}
hist(as.numeric(dataTraining$classe), xlab="Classe", main="Training Data Set", axes=FALSE)
axis(1, at = c(1,2,3,4,5), labels = c("A","B","C","D","E"))
axis(2)
```

### Data Cleaning### 

Remove timestamps, names, etc that won't help predict
```{r}
r <- c(1:5)
dataTraining <- dataTraining[-r]
```

Remove columns with more than 90% NAs
```{r}
dataTraining <- dataTraining[,colSums(is.na(dataTraining)) < (nrow(dataTraining) * .9)]
```

Remove attributes that won't help predict
```{r}
nzv <- nearZeroVar(dataTraining)
dataTraining <- dataTraining[-nzv]
```

### Create model### 
The random forest algorithm is used to construct the model

```{r}
library(randomForest)
model <- randomForest(classe~.,data=dataTraining)
model
```

The model estimates the error rate at 0.49% 

### Evaluate testing data### 
```{r}
p <- predict(model, dataTesting)
cm <- confusionMatrix(dataTesting$classe, p);
cm
```
The prediction yields an accuracy rate of 99.5%, which means an error rate of 0.5%, so almost identical to the predicted error rate of 0.49%

### Predict values for pmlTest data ### 
```{r}
p2 <- predict(model, pmlTest)
p2
```

